### 2021-08-25 MResProject/Latex/Draft.tex
CMEECourseWork

\documentclass[11pt]{article}

\usepackage{setspace}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=red,citecolor=black]{hyperref}
\usepackage{lineno}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}
\usepackage{floatrow}
\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{geometry}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{natbibspacing}
\usepackage[toc,page]{appendix}
\usepackage{makecell}

% Metagenomic DNA discoveries after sequencing everything (bacteria, parasites, food, gut) inside a bee
\title{Optimize sequencing depth for shotgun metagenomics of pollination system by rarefaction, using a modular profiling pipeline}
\author{Cong Liu}
\date{2021.8}

\linespread{1.5}
\geometry{left=2cm,right=2cm,top=2cm,bottom=2cm}

\begin{document}

\bibliographystyle{plainnat}
\setcitestyle{round}

  \maketitle
  \vspace*{\fill}
  \begin{center}
    A thesis submitted in partial fulfilment of the requirements for the degree of Master of Research at Imperial College London
    \newline
    Submitted for the MRes in Computational Methods in Ecology and Evolution
  \end{center}

  \newpage
  Declaration:
  \newline
  All datasets used in this project are provided by Dr. Peter Graystock and the project was conducted under his supervision. 
  I am responsible for the development and conduction of the analyses presented.

  \newpage
  \linenumbers
  \begin{abstract}
    Bee gut microbiome plays a crucial role in host health, involving in food digestion, pathogen defence and chemical detoxification. 
    Vast relevant investigations use amplicon sequencing to explore bacterial diversity, while emergence of shotgun metagenomics offers unique advantages by capturing diversity of multiple taxonomic clades and providing information on function potentiality. 
    However, utilization of shotgun metagenomics is hindered by complexity of data analysis and high cost of sequencing. 
    % However, bioinformatics for metagenomic data analysis is complex, and the results may be compromised by insufficient sequencing depth. 
    Here, an integrated pipeline combining assembly-dependent and -independent methods was introduced for taxonomic and functional profiling of shotgun metagenomic data, and a framework of rarefaction and multimodel inference was constructed for optimizing sequencing depth. 
    Both the pipeline and the framework were used for analysis of deep-sequencing datasets ($2 \times 150 $bp read pairs) from of honey bees, bumble bees and flower washes. 
    The integrated pipeline illustrated taxon composition and metabolic potentiality of the metagenomes, and provided significant improvements in species identification. 
    For species diversity estimation, about 40 and 43 million read pairs would be sufficient for metagenomic datasets from honey bees and bumble bees, respectively. 
    If low leverage is given to rare species, shallower sequencing can be adopted. 
    Function profiling is much less stable to sequencing depth than species content. 
    Additional functional gene clusters (FGCs) were still being discovered at original sequencing depth of all samples. 
    % Sequencing depth can be optimized for species diversity estimation especially when little emphasis is given to rare species, while functional profiling requires deeper sequencing than taxon profiling. 
    % Rarefaction and model fitting was conducted to compute expected species/function diversity (Hill numbers of order 0, 1, 2) returned by given sequencing effort in order to optimize sequencing depth, balancing sequencing effort and capability of estimating diversity reliably. 
    % As a result, functional profiling was demanding about sequencing effort. 
    % None of datasets involved was sufficient for providing almost complete coverage of function diversity. 
    % As for species diversity, the average optimal sequencing depth for reliable estimation of Hill number of order 0, 1 and 2 for honey bee samples are 40.33 million (12.10 Gbp), 18.57 million (5.57 Gbp) and 17.45 million (5.24 Gbp), respectively; and for bumble bee samples, the averages are 42.49 million (12.75 Gbp), 40.33 million (12.10 Gbp) and 24.77 (7.43 Gbp) million. 
    % For flower surface, more than 0.88 million (0.26 Gbp) sequencing depth is needed. 
    Overall, this project provides guidelines for defining a balance between sequencing cost and acquirement of reliable results for investigations of pollination system. 
    Similar studies for other host species are recommended before undertaking metagenomic projects involving a large number of samples. 
    % The integrated pipeline and the framework for rarefaction analysis can also be applied for researches on other environments. 
  \end{abstract}

  \section{Introduction}
  Pollinators such as bees are crucial to maintain global food security and provide stability of natural systems \citep{hristov2020significance,bansch2021crop,khalifa2021overview} but their populations are facing declines \citep{brown2009conservation,hristov2020factors,cheng2021ecology,zattara2021worldwide}. 
  % The health of bees is a crucial issue that has received increasing concern \citep{amiri2020using}. 
  % As pollinators of numerous plants, bees play an important role in the stability of natural and agricultural systems \citep{hristov2020significance,bansch2021crop,khalifa2021overview}.
  % foldesi2021larger 
  % Diverse pathogens and environmental factors threaten bee health and are contributing to their declines. 
  % The trend of bee decline has been widely observed at a global scale and attributed to multiple factors including 
  A combination of stressors including parasites, pesticide exposure, invasive species, habitat loss and climate change is contributing to declines of bee populations. \citep{brown2009conservation,hristov2020factors,cheng2021ecology,zattara2021worldwide}.
  \newline
  % Bees are associated with a community of microorganisms influencing host health \citep{engel2016bee,raymann2018role}. 
  Within the gut of bees, stable microbial communities (microbiomes) play crucial roles in food digestion, parasite defence and chemical detoxification \citep{moran2015genomics,engel2016bee,raymann2018role}. 
  They help mediate food digestion processes including 
  % plant-produced polysaccharide including pectin and hemicellulose \citep{zheng2019division}
  polysaccharides breakdown \citep{zheng2019division}, sucrose hydrolysis \citep{engel2012functional,lee2015saccharide}, 
  % a predominant component of nectar \citep{nicolson2007nectar} 
  and mannose metabolism \citep{engel2012functional,lee2015saccharide}. 
  % which presents in nectar of \textit{Tilia} \citep{adler2000ecological} and is poisonous to honey bees \citep{de1986mechanism}. 
  Bee microbiomes also provide protection against pathogens including \textit{Crithidia}, 
  % a protozoan genus exclusively parasitising arthropods 
  \citep{koch2011socially,cariveau2014variation}, \textit{Paenibacillus larvae}, 
  % a bacteria causing lethal American foulbrood in honey bees 
  \citep{ebeling2016biology,forsgren2010novel} and \textit{Nosema sp.} 
  % a genus of fungal pathogens 
  \citep{cariveau2014variation,maes2016diet}, 
  and are involved in resistance to both metal and metalloid toxins including cadmium \citep{rothman2019cadmium}, copper \citep{rothman2020direct} and selenate \citep{rothman2019bumble}. 
  \newline
  Amplicon sequencing, in which a species-specific barcode region is amplified and sequenced, is a powerful and vastly used method for investigations of microbiome compositions \citep{abdelfattah2018metabarcoding}.  
  In bee microbiome investigations, it is used to illustrate the taxonomic diversity of bee-associated bacteria and fungi (\textit{e.g.} \cite{geldert2021dietary,wang2021gut,powell9field,kapheim2021composition}). 
  However, bees visit numerous niches during environment exploration and foraging activities, and can get contact with diverse eDNA signatures, which provide insight into pollinator ecology by reflecting interactions between bees and other organisms including bacteria, fungi, plants, arthropods and viruses \citep{bovo2018shotgun,ribani2020honey,bovo2020shotgun,matsuzawa2020application}. 
  It is difficult to explore this diversity via amplicon sequencing because it only captures a fraction of the whole community since analysis of different taxonomic groups is based on different barcode regions, \textit{e.g.} 16S ribosomal RNA (rRNA) for bacteria \citep{hayashi2002phylogenetic,eckburg2005diversity}, internal transcribed spacer (ITS) for fungi \citep{nilsson2008intraspecific}, cytochrome c oxidase subunit I (COI) for Animalia \citep{hebert2003barcoding} 
  and plastid genes for plants \citep{group2009dna}. 
  As a result, amplicon sequencing only captures taxon diversity within a certain clade. 
  Besides, it is difficult to illustrate function potentiality of bee-associated microbiome using amplicon sequencing since it does not provide information on content of functional gene clusters (FGCs), \textit{i.e.} aggregates of genes with same function. 
  As a result, functional capacity needs to be inferred based on taxon composition. 
  Amplicon-based function predictors such as Tax4Fun \citep{asshauer2015tax4fun} and PICRUSt \citep{douglas2018predicting} predict functional capacity based on pre-sequenced genomes without taking gene content variation within taxa. 
  However, bee bacterial symbionts are diversified at strain level \citep{engel2012functional,powell2016strain,ellegaard2020vast} and bacterial strains are often highly variable in gene content \citep{cordero2014explaining,brockhurst2019ecology}. 
  As a result, amplicon-based inference of functional capacity of bee microbiome may not be reliable.
  % Host-associated microbiome plays an important role in host homeostasis and metagenomics boosted by high throughput sequencing provides powerful tools for understanding its composition, function and dynamics. 
  % For example, amplicon sequencing is a powerful tool and has revealed incredible understanding in microbiome \citep{eckburg2005diversity,galloway2020tools,new2020metagenomics}. 
  % However, it only focuses on a small part of the whole microbiome. 
  % Shotgun sequencing of microbiome, or metagenomics, unselectively captures DNA in a sample and is capable of providing comprehensive inventories of taxa and functional gene clusters (FGCs) with high resolution \citep{quince2017shotgun,new2020metagenomics}.
  \newline
  Shotgun metagenomics provides a solution for microbiome investigations to overcome drawbacks of amplicon sequencing. 
  By capturing and sequencing DNA fragments unselectively, shotgun metagenomics is capable of providing comprehensive inventories of taxa and FGCs \citep{quince2017shotgun,new2020metagenomics,galloway2020tools}. 
  However, utilization of shotgun metagenomics is hindered by challenges in bioinformatics. 
  The goal of metagenomics is typically to provide a taxonomic and functional profile of the microbiome, and there is not a gold standard for performing metagenomic data analysis. 
  % bioinformatics of metagenome. 
  Generally, one of the first steps in metagenomic analysis is assembling short reads into long contigs, which can help improve accuracy of metagenomic annotation \citep{wommack2008metagenomics,carr2014comparative,tran2020assembling} and is necessary for 
  % tasks including recovery of uncultured genomes \citep{garza2015cultured,chen2020accurate} and 
  discovery of novel taxa and genes \citep{culligan2014metagenomics,youngblut2020large}. 
  However, metagenome assembly is complex, compromised by fragmental assembly, chimaeras \citep{mikheenko2016metaquast} and loss of taxon/function diversity due to unassembled reads \citep{vollmers2017comparing,ayling2020new}. 
  Probably because of these shortcomings, assembly is sometimes skipped and short reads are directly proceeded for annotation \citep{tringe2005comparative,abubucker2012metabolic,vermote2018amplicon,bovo2018shotgun}, although the accuracy can be compromised due to low information load of short reads \citep{wommack2008metagenomics,carr2014comparative,tran2020assembling}. 
  A combination of both assembly-dependent and -free methods could overcome the complexity and improving accuracy of metagenomic profiling \citep{becker2020modular}.
  % complexity of metagenomic assembly and inaccuracy of short read annotation \citep{becker2020modular}. 
  \newline
  A remaining challenge of metagenomics is the determination of a sequencing depth that provides reliable estimation of taxon/FGC diversity without overspending. 
  It is recommended to retrieve as many reads as possible \citep{quince2017shotgun} since insufficient sequencing causes compromise in metagenome profiling \citep{cattonaro2018you,zaheer2018impact,pereira2019impact,gweon2019impact}. 
  However, deep metagenomic sequencing is expensive, which hinders its utilization, especially in large-scale projects. 
  Currently, there are few published guidelines for the sufficient sequencing depth of a given environment or study type in order to reach a trade-off between sequencing effort and reliable output. 
  \newline
  In order to balance sequencing cost and reliable estimation of taxon/FGC diversity, expected diversity represented by given sequencing depth need to be computed. 
  Since a metagenomic dataset can be viewed as a random sample of an assemblage of genomic sequences, and profiling is the process by which reads are assigned to taxa or FGCs, the relationship between sequencing depth and diversity can be illustrated by rarefaction (randomly subsampling the original dataset without replacement) and quantified by model fitting, if the original dataset can provide an almost complete inventory of taxa/FGCs \citep{heck1975explicit,hortal2005ed,gomez2014using,hughes2021sampling}. 
  The sequencing depth is sufficient for reliable estimation of taxon/FGC diversity if and only if the slope of the rarefaction curve is small \citep{hortal2005ed,chao2012coverage,roswell2021conceptual}. 
  % The determination of sequencing depth can be conducted by rarefaction analysis, a method widely used in traditional ecology \citep{heck1975explicit,hortal2005ed,gomez2014using,hughes2021sampling}. 
  % In field based surveys, a random sample of individuals is drawn from a community and assigned to species. 
  % Then a rarefaction curve illustrating expected biodiversity represented by given sampling effort (often measured by number of captured individuals) is generated by random subsampling the original sample without replacement, and quantified by model fitting \citep{hughes2005application,gotelli2011estimating}. 
  % The slope of the rarefaction curve represents the expected rise of the curve if one more individual is captured. 
  % The sample is nearly complete if and only if the final slope is small, and the point at which the slope of rarefaction curve falls to a cut-off value represents the minimal sampling effort required for assessment of biodiversity \citep{hortal2005ed,chao2012coverage,gomez2014using,roswell2021conceptual}. 
  % This framework for determination of sampling effort can be used in shotgun metagenomics, since a metagenomic dataset can be viewed as a random sample of an assemblage of genomic sequences, and 
  % profiling is the process by which reads are assigned to taxa or FGCs. 
  % The concept corresponding to sampling effort is sequencing depth in metagenomics, which is measured by read number.
  % \newline 
  % In macroecology, comprehensive analytical frameworks are available for computing expected biodiversity returned by given sample sizes using a reference sample \citep{heck1975explicit,chao2012coverage,chao2014rarefaction}. 
  % % However, analytical approaches derived for computing rarefaction curve in macroecology \citep{heck1975explicit,chao2012coverage,chao2014rarefaction} 
  % However, they are not suitable for metagenomics because 
  % % they are based on 
  % of the assumptions that every captured individual can be assuredly assigned to a species, and the present of rare species is as reliable as that of abundant species. 
  % In shotgun metagenomics, reads may be unannotated or annotated incorrectly due to low sequencing quality, host contamination, similar structures in genomes of different species, non-coding regions in genomes and limited sensitivity of profiling pipeline. 
  % As a result, taxa or FGCs of low relative abundance are more likely to be false positive than that of high abundance.
  \newline
  In this project, I utilised metagenomic datasets ($2 \times 150 $bp) from three environmental types: the gut of honey bees (\textit{Apis mellifera}), the gut of common North American bumble bees (\textit{Bombus impatiens}) and the surface of a wild flower (\textit{Erigeron annuus}). 
  I aimed to (1) develop an integrated pipeline combining assembly-dependent and -independent methods to deliver improved taxon annotation of sequencing data and (2) optimize sequencing depth to balance sequencing cost and requirement for reliable analysis of microbial species diversity, or the description of their functional diversity. 
  The integrated pipeline provided improvement in species identification compared with the assembly-dependent method. 
  Rarefaction analysis showed that about 40/43 million clean read pairs would be a suitable compromise for sequencing honey/bumble bee samples and species diversity detection. 
  Shallower sequencing can be adopted with reduced emphasis on rare species. 
  Functional profiling is more demanding about sequencing depth than species profiling. 
  Significant accumulation of FGCs was observed at final points of rarefaction curves from all samples.  
  These results provide guidelines for defining a balance between sequencing cost and obtaining reliable taxonomic/functional profiles for metagenomic investigations of pollination system. 
  Similar studies for other host species are recommended before undertaking metagenomic projects with big sample size. 
  % In the integrated pipeline (Figure \ref{Pipeline}), 
  % assembly-dependent taxon profiling is conducted after quality filtering and removing host contamination. 
  % To address high negative rate of assembly-dependent search, a reference database of genomes from species present in the assembly is constructed and used for filtering non-host reads. 
  % Reads not recruited by the reference database are subjected to assembly-free taxonomic search. 
  % This integrated pipeline was used to profile metagenomic datasets involved in this project. 
  % % It was shown that the construction of reference database and assembly-free search helped identified species not represented by the assembly. 
  % Then in order to estimate optimal sequencing depth for reliable representing taxonomic/functional diversity of metagenomes, I simulated different sequencing depth by rarefaction and profiled subsampled datasets using the integrated pipeline, generating inventories of species and FGCs represented by Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) \citep{kanehisa2000kegg}. 
  % Diversity of species and KOs was measured by Hill numbers of order 0, 1 and 2. 
  % Sequencing depth was measured by number of clean read pairs (150 bp read length). 
  % % The higher the order is, the more sensitive to relative abundances Hill number is. 
  % The relationship between Hill numbers and sequencing depth was quantified by fitting and averaging asymptotic species accumulation models, and estimation of optimal sequencing depth was given by the point where the slope of rarefaction curve drops to cut-off values. 
  % % Sample would be excluded from optimization of sequencing depth if the original dataset was not sufficient for providing an almost complete inventory of species or KOs, which was characterized by small final slope of rarefaction curve for species or KO richness (Hill number of order 0). 
  % % As a result, when measuring species diversity by Hill number of order 0, 1 and 2, the average optimal sequencing depth for honey bee samples are 40.33 million, 18.57 million and 17.45 million; for bumble bee samples, the averages are 42.49 million, 40.33 million and 24.77 million; and for flower surface, 0.88 million sequencing depth is not sufficient for providing an almost complete inventory of species, indicating higher sequencing depth is needed. 
  % % As for functional diversity, no datasets involved in this project is big enough for providing an almost complete inventory of KOs, indicating the sequencing depth need to be higher than the maximum of each sample type, \textit{i.e.} 45.06 million for honey bees, 53.61 million for bumble bees and 0.88 million for flower surface. 
  % % The integrated pipeline and the framework for rarefaction analysis can also be applied for investigations of microbiomes from other environments, and the estimations of optimal sequencing depth provide guidelines for cost-effective utilization of shotgun metagenomics in investigations of pollination system.

  \section{Materials and Methods}
    \subsection{Samples, DNA extraction and sequencing}
    Samples include four honey bees (two from the same hive (hive 13), one from a hive in the same apiary (hive 15) and one caught foraging), three bumble bees from commercially supplied (Biobest) bumblebee colonies (two from the same colony), and one buffer wash of a wild flower (\textit{Erigeron annuus}).  
    \newline
    DNA extraction was performed as in \cite{graystock2020dominant}, followed by library preparation using a low template protocol with Illumina Nextera Library Prep kits. Briefly, this involved tagmentation into fragments of 300 bases before eight samples of 10ng were pooled together and sent to Beijing Genomics Institute (BGI) for further quality control and sequencing using a full lane in the X-ten platform.

    \subsection{Metagenomic profiling using integrated pipeline}
      \subsubsection{Integrated pipeline}

      \begin{figure}[H]
        \centering
        \includegraphics[width=\textwidth]{../Figures/FigurePipeline.pdf}
        \caption{An overview of the integrated pipeline. 
        The pipeline is separated into 5 modules, indicated by blue boxes. 
        Green boxes indicate dataset of reads, while yellow boxes indicate database of genomes. 
        Black arrows indicate the input and output of each step and the numbers on them indicate the order that each step is utilized.}
        \label{Pipeline}
      \end{figure}

      I designed an integrated pipeline for shotgun metagenomic profiling, \textit{i.e.} assigning reads to taxa or FGCs. 
      It is separated into five modules (Figure \ref{Pipeline}).
      \newline
      In quality control and host removing, raw sequencing data quality is checked using FastQC v.0.11.5 \citep{andrews2010fastqc} and filtered by Trimmomatic v.0.39 \citep{bolger2014trimmomatic}. 
      Then clean reads are mapped to host genome using Bowtie2 v.2.4.2 \citep{langmead2012fast} and non-host reads are extracted by SAMtools v.1.11 \citep{li2009sequence}. 
      \newline
      The non-host reads are subject to the module of assembly-dependent species identification. 
      \textit{De novo} assembly is conducted by SPAdes v.3.15.2 \citep{prjibelski2020using}. 
      Assembled contigs are aligned to NCBI non-redundant (nr) database by DIAMOND v.2.0.7.145 \citep{buchfink2015fast}, and assigned to taxa by MEGAN6 \citep{huson2007megan}. 
      \newline
      Then fragment recruitment is conducted to filter reads mapped to species represented by the assembly. 
      A reference database is constructed and it comprises reference genome dataset, \textit{i.e.} genomic sequences in FASTA format and corresponding genome annotation in general feature format (gff). 
      For each species represented by assembly, its reference genome dataset, if available, is downloaded from NCBI using its \textit{datasets} command-line tool and added to the reference database. 
      Then non-host reads are mapped to the reference database by Bowtie2, and unmapped reads are extracted by SAMtools. 
      \newline
      Reads not recruited by the reference database are subjected to assembly-independent species identification. 
      They are aligned to NCBI nr database through DIAMOND and assigned to taxa by MEGAN6.
      \newline
      Finally, functional annotation is conducted by EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      It takes coding sequences (CDSs) of genomes in the reference database and reads subject to assembly-independent species identification as input and assigns them to Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) \citep{kanehisa2000kegg}.
      \newline
      The integrated pipeline was used for analysing metagenomic datasets involved in this study. 
      Details in the pipeline and parameter settings of each module are described in Supplementary \ref{ParameterSetting}.

      \subsubsection{Taxon/function quantification and metabolic pathway reconstruction}
      After profiling, identified species and KOs were quantified by calculating relative sequence abundance, \textit{i.e.} proportion of reads assigned to a species/KO in all reads annotated. 
      For species without available reference genomes, their abundances were calculated using reads assigned to them in assembly-independent search. 
      As for taxa with available reference genomes, they may be identified in both assembly-dependent and -independent search due to strain-specific genomic structures that are not present in reference genomes. 
      Their abundances were calculated by summating number of reads that (1) mapped to coding sequences (CDSs) of reference genomes and (2) assigned to them in assembly-independent search. 
      Reads mapped to non-coding regions were not taken into consideration in order to avoid overestimation since the assembly-independent search was based on aligning reads to nr database, which is composed of proteins. 
      As for KO quantification, CDSs with zero coverage were excluded. 
      Abundances of KOs were calculated by summating number of reads that (1) mapped to CDSs assigned to KOs and (2) assigned to KOs directly. 
      Extraction of CDSs and calculation of their coverage were conducted by BEDtools v.2.30.0 \citep{quinlan2010bedtools}.
      % \newline
      % In order to investigate metabolic potentialities of metagenome, CDSs with non-zero coverage from reference genomes and reads subject to assembly-independent search were functionally annotated using EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
      % It searches query sequences  
      % Identified functional categories were quantified by proportion of reads assigned to a category in all reads annotated (relative sequence abundance). 
      % The number of reads assigned to a functional category equals the sum of reads (1) assigned to the category directly and (2) mapped to CDSs assigned to the category.
      \newline
      Metabolic pathways were inferred based on KOs. 
      Reads assigned to plants and arthropods were not included since they were unlikely to represent living organisms. 
      MinPath v.1.6 was used for pathway inference \citep{ye2009parsimony}. 
      It finds a minimal set of KEGG pathways that can explain all KOs provided as input.

    \subsection{Estimation of optimal sequencing depth required for metagenomic profiling}
    % The impact of sequencing depth on taxon/function diversity of metagenome was simulated by rarefaction. 
    % Diversity was measured by Hill numbers, which are calculated from inventory of relative abundances. 
    % For taxon diversity, the inventory was generated by assigning reads to species; while for function diversity, it was obtained by assigning reads to KOs. 
    % Then the rarefaction curve plotting Hill number against sequencing depth was computed, and optimal sequencing depth required for reliable estimation of diversity were given by the point where the slope of rarefaction curve drops to a cut-off value.

      \subsubsection{Simulating different sequencing depth by subsampling and diversity measured using Hill numbers}
      Species/KO inventories obtained from different sequencing depth was simulated by rarefaction. 
      % \textit{i.e.} randomly taking a proportion of the original dataset without replacement. 
      Since the ratio between numbers of raw and clean reads is dependent on sequencing process and is not influenced by sample type, sequencing depth here refers to number of clean read pairs to exclude variance caused by different proportion of low quality reads in samples of same type. 
      Besides, the expected ratio between host and non-host reads in a metagenomic dataset is dependent on the DNA sample and not impacted significantly by sequencing depth. 
      Thus the proportion of non-host reads in each simulation is expected to be the same with that in the original dataset. 
      \newline
      Based on these considerations, I randomly subsampled non-host dataset of each sample, taking 10\%-100\% of read pairs at interval of 10\% by \textit{reformat.sh} script of BBmap v.38.90 \citep{bushnell2014bbmap}, and profiled subsampled datasets by the integrated pipeline (Figure \ref{Pipeline}). 
      The sequencing depth of each subsampled dataset equals number of subsampled non-host read pairs divided by ratio between non-host and clean read pairs. 
      Thus, each subsampled dataset of non-host reads is corresponded to an imaginary dataset of clean reads, whose proportion of non-host reads is the same with that of the original metagenomic dataset. 
      \newline
      After profiling subsampled datasets, species/KO diversity was measured by Hill number, a parametric family of diversity indexes differing by order $q$ \citep{hill1973diversity,ma2018measuring,roswell2021conceptual}. 
      Hill numbers provide unique advantages over other diversity indexes \citep{chao2014unifying,chao2014rarefaction,alberdi2019guide,roswell2021conceptual}. 
      First, Hill numbers follow a replication principle: if a proportion of categories in an inventory was removed randomly, all Hill numbers are expected to decrease by that proportion. 
      Second, the sensitivity of Hill numbers to relative abundances can be modulated by order $q$. 
      % Third, all Hill numbers are pressed in the units of effective numbers of species: the number of equally abundant species in an imaginary inventory with the same diversity. 
      Third, widely used diversity indexes including species richness, Shannon index and Simpson index, can be converted to Hill numbers through algebraic transformations. 
      \newline
      Hill number of order $q$ is defined as Equation \ref{HillNumbers} \citep{hill1973diversity}.
      \begin{equation}
        D^{(q)} = (\sum_{i}(p_i)^q)^{\frac{1}{1-q}}
        \label{HillNumbers}
      \end{equation}
      $p_i$ represents the relative abundance of $i$th species/KO. 
      When $q = 0$, abundances are not taken into consideration and $D^{(0)}$ equals species/KO richness. 
      When $q = 1$, emphasis is given to species/KOs with general abundances. 
      Hill number is defined as the limit of Equation \ref{HillNumbers} as $q$ tends to 1 and equals the exponential of Shannon index (\ref{HillShannon}).  
      \begin{equation}
        D^{(1)} = e^{-\sum_{i}p_ilogp_i}
        \label{HillShannon}
      \end{equation}
      When $q = 2$, high leverage is provided to abundant species/KOs and Hill number equals the inverse of Simpson index (Equation \ref{HillSimpson}).
      \begin{equation}
        D^{(2)} = \frac{1}{\sum_{i}(p_i)^2}
        \label{HillSimpson}
      \end{equation}

      \subsubsection{Quantification of relationship between sequencing depth and Hill numbers by fitting rarefaction curves}
      Hill number of order $q$ (Equation \ref{HillNumbers}) measures diversity of an inventory as the number of equally abundant categories in an imaginary inventory with the same diversity \citep{chao2014unifying,roswell2021conceptual}. 
      Order $q$ determines leverage given to abundant categories. 
      All Hill numbers behave in the following way: removing a proportion of categories in an inventory randomly is expected to cause decrease of Hill numbers in that proportion \citep{roswell2021conceptual}. 
      Thus, it can be hypothesized that as sequencing depth (number of clean read pairs) increases, the detection of novel species/KOs leads to increase of Hill numbers, and when sequencing depth is so big that all species/KOs present in the metagenomic DNA sample have been detected, Hill numbers level off. 
      Such a relationship can be fitted by asymptotic species accumulation models.
      \newline
      Let Hill number of order $q$ (Equation \ref{HillNumbers}) be a function of sequencing depth $x$, which takes million read pairs as the unit. 
      This function was fitted using a multimodel inference method. 
      First, a total of five candidate models (Table \ref{candidate_models}) were fitted to rarefaction curves which plots Hill numbers against sequencing depth. 
      R package \textit{minpack.lm} v.1.2.1, which employs Levenberg-Marquardt nonlinear least-square algorithm, was used for model fitting. 
      Then small sample unbiased Akaike information criterion (AICc) \citep{anderson2007model} of each candidate model was calculated (Equation \ref{AICc}):
      \begin{equation}
        AICc = -2L+2k+\frac{2k(k+1)}{(n-k-1)}
        \label{AICc}
      \end{equation}
      where $n$ is number of observed data points ($n = 10$ in this study), $k$ is the number of fitted coefficients, and $L$ is maximized log-likelihood, given by Equation \ref{likelihood}.
      \begin{equation}
        L = -0.5nlog(\frac{Rss}{n})
        \label{likelihood}
      \end{equation}
      $Rss$ represents residual sum of squares.
      \newline
      Then model averaging was conducted. First, differences of AICc scores between $i$th candidate models and the model with lowest AICc value were calculated using Equation \ref{Delta}.
      \begin{equation}
        \Delta_i = AICc_i-AICc_{min}
        \label{Delta}
      \end{equation}
      $AICc_i$ is the AICc score of $i$th plausible model and $AICc_{min}$ is the lowest AICc score among all candidate models. 
      The Akaike weight of $i$th model is given by Equation \ref{weight} \citep{anderson2007model}.
      \begin{equation}
        w_i = \frac{e^{(-0.5\Delta_i)}}{\sum_{i}e^{(-0.5\Delta_i)}}
        \label{weight}
      \end{equation}
      Denote $i$th candidate model by $D^{(q)}_i = D^{(q)}_i(x)$, the averaged model is given by Equation \ref{average}.
      \begin{equation}
        D^{(q)}(x) = \sum_{i}w_iD^{(q)}_i(x)
        \label{average}
      \end{equation}
      The slope of rarefaction curve was calculated by first derivative of averaged model (Equation \ref{average_derivation}). 
      It reflects the increase rate of the curve.
      \begin{equation}
        \frac{dD^{q}}{dx} = \sum_{i}w_i\frac{dD^{(q)}_i}{dx}
        \label{average_derivation}
      \end{equation}
      The asymptote of rarefaction curve as sequencing depth tends to infinity is given by Equation \ref{asymptote}. 
      It provides an estimation of the total diversity and is comparable among metagenomic DNA samples with different sequencing depth. 
      % \citep{lamas1991preliminary,soberonm1993use,hortal2004butterfly,jimenez2005determining,hortal2006evaluating}. 
      However, the accuracy of asymptotic estimators is controversial \citep{colwell1994estimating,chazdon1998statistical,jimenez2006comparing,hortal2006evaluating}.
      \begin{equation}
        \lim_{x \to +\infty}D^{(q)}(x) = \sum_{i}w_i\lim_{x \to +\infty}D^{(q)}_i(x)
        \label{asymptote}
      \end{equation}

      \subsubsection{Estimating optimal sequencing depth using rarefaction curves} 
      Optimal sequencing depth is defined as the point at which diversity starts to level off as sequencing depth increases, and its precise estimation via rarefaction is based on the assumption that the original dataset is sufficient for detection of almost all species/KOs present. 
      % providing a reliable inventory of species/KOs present, \textit{i.e.} almost all species/KOs present in the DNA sample are detected. 
      This assumption can be verified by looking at the rarefaction curve that plots species/KO richness (Hill number of order 0) against sequencing depth. 
      The original dataset is sufficient for providing a reliable inventory if and only if the rarefaction curve of richness is characterized by a small final slope \citep{heck1975explicit,hortal2005ed,chao2012coverage}. 
      Then an estimation of optimal sequencing depth is provided by the point at which the slope of rarefaction curve decreases to a given cut-off value \citep{hortal2005ed,gomez2014using}. 
      % \newline 
      % This method is on the analogy of that used in macroecology to determine whether more sampling effort is needed to illustrate biodiversity of a community. 
      % In macroecological researches, rarefaction curve plots detected biodiversity against sampling effort, which is often measured by number of captured individuals. 
      % Precise quantification of the relationship between biodiversity and sampling effort via rarefaction assumes the original sample is nearly complete \citep{heck1975explicit,gotelli2011estimating}. 
      % This assumption can be verified by looking at the slope of rarefaction curve, which is the expected rise of the curve if one more individual is captured. 
      % The sample is nearly complete if and only if the final slope is small\citep{heck1975explicit,chao2012coverage,roswell2021conceptual}. 
      % The point at which the slope of rarefaction curve falls to a cut-off value represents the minimal sampling effort required for assessment of biodiversity \citep{hortal2005ed}. 
      % The concept corresponding to sampling effort is sequencing depth in metagenomics, which is measured by number of read pairs. 
      % Thus, the insights provided by rarefaction analysis in macroecology is transferrable to metagenomics.
      % \newline 
      % However, analytical approaches derived for computing rarefaction curve in macroecology \citep{chao2012coverage,chao2014rarefaction} are not suitable for metagenomics because they are based on the assumption that every captured individual can be assuredly assigned to a species, which is unreasonable for shotgun metagenomics. 
      % In shotgun metagenomics, reads may not be assigned to any species or function category for multiple reasons, including sequencing error, limited sensitivity of homology search, similar structures in genomes of different species and the present non-coding regions.

  \section{Results}

    \subsection{Sequence reads}
    Eight samples (four honey bees, three bumble bees and one flower eDNA) were sequenced. 
    The quality reports of raw reads showed low-quality 3'-end (Figure \ref{QC}a), uneven base content in 5'-end (Figure \ref{QC}b) and the present of adaptors (Figure \ref{QC}c). 
    The quality control procedure covered these aspects and improved data quality (Figure \ref{QC}d, \ref{QC}e and \ref{QC}f).
    \newline
    After quality control, read pairs aligned to host genome were removed. 
    Table \ref{data_statistics} reports numbers of raw, clean and non-host read pairs. 
    Honey bee sample \textit{Bee\_Amellifera\_13\_1} was filtered from further analysis since its low raw read pair number (1.10 million compared to the other samples at ~59 million) is suggestive of a poor quality sample.  
    After quality control, 62.08\%-76.52\% of raw read pairs were retained for these three honey bee samples. 
    Then a different proportion of non-host read pairs (29.52\%-86.19\%) were retained from each sample. 
    As for bumble bee samples, about 58 million raw read pairs were obtained for each sample and 82.1\%-83.8\% were retained after quality control. 
    After host removing, 7.35\%-10.08\% of clean reads were retained.  
    For flower eDNA sample, 1.44 million raw read pairs were obtained and 61.15\% of them were retained. 
    Host removing was not conducted because it is a sample of eDNA washed from flower surface.

    % \subsection{Characterization of species composition}
    \subsection{Application of integrated pipeline}
    The integrated pipeline was used to profile metagenomic datasets from pollination system, illustrating its capacity in presenting taxon composition and functional potentiality of microbiome. 
    Although samples are different in proportion of host contamination (Table \ref{data_statistics}, Figure \ref{GroupAbundance}a), diverse communities composed of multiple taxonomic clades were identified (Figure \ref{GroupAbundance}b). 
    Most species identified are common in pollination system. 
    Their present indicate that samples were of good quality and representative for microbiomes from pollination system. 
    % All seven samples were dominated by bacterial and/or arthropod species (Figure \ref{GroupAbundance}b). 
    % Honey bee samples \textit{Bee\_Amellifers\_hv13\_2} and \textit{Bee\_Amellifera\_wild\_1} were dominated by bacteria (approximate 90\% relative abundance), while in \textit{Bee\_Amellifera\_hv15\_1}, arthropods, bacteria and viruses accounted for most annotated reads. 
    % As for bumble bee samples, they were all dominated by bacteria and arthropods. 
    % For the flower eDNA sample, arthropods were the most dominant and bacteria were the second.

      \begin{figure}[H]
        \begin{subfigure}
          \centering
          \textbf{a}
        \end{subfigure}
        \hspace{0.5\linewidth}
        \begin{subfigure}
          \centering
          \textbf{b}
        \end{subfigure}
        \newline
        \hfill
        \begin{subfigure}
          \centering
          \includegraphics[width=0.48\linewidth,height=0.48\linewidth]{../Figures/NonHostRatio.pdf}
        \end{subfigure}
        \hfill
        \begin{subfigure}
          \centering
          \includegraphics[width=0.48\linewidth,height=0.48\linewidth]{../Figures/1_species_RelAbundance_GroupAbundance.pdf}
        \end{subfigure}

        % \begin{subfigure}
        %   \centering
        %   \textbf{a}
        % \end{subfigure}
        % \newline
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.5\textwidth]{../Figures/NonHostRatio.pdf}
        % \end{subfigure}
        % \vfill
        % \newline
        % \begin{subfigure}
        %   \centering
        %   \textbf{b}
        % \end{subfigure}
        % \newline
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.5\textwidth]{../Figures/1_species_RelAbundance_GroupAbundance.pdf}
        % \end{subfigure}
        % \vfill
        \caption{The number of host and non-host reads in each sample (a) and the relative abundance of species under six taxonomic groups (b): superkingdom Viruses, superkingdom Bacteria, kingdom Viridiplantae (plants), kingdom Fungi, phylum Arthropoda and others (species that are not in the other five groups).}
        \label{GroupAbundance}
        \end{figure}

    The presence of known core members of bee-associated bacterial community including species within \textit{Bifidobacterium}, \textit{Frischella}, \textit{Gilliamella}, \textit{Snodgrassella}, \textit{Lactobacillus}, \textit{Apilactobacillus} and \textit{Bombilactobacillus} (Figure \ref{BacteriaHeatmap}) \citep{koch2011bacterial,moran2015genomics,kwong2017dynamic,zheng2020taxonomic} suggests good sample quality. 
    As for other abundant bacteria, \textit{Fructobacillus sp.} are often found in fructose-rich environments like flowers \citep{endo2014genus}; \textit{Bartonella apis} is related to animal pathogens \citep{kevsnerova2016bartonella} and is widespread in honey bee workers \citep{raymann2018role}; \textit{Candidatus Schmidhempelia bombi} is a known uncultured symbiont of \textit{Bombus impatiens} \citep{martinson2014genomic}. 
    It should be noted that some typical bee-associated bacteria were also found on the flower, including \textit{Bartonella apis}, \textit{Bifidobacterium asteroides}, \textit{Bombilactobacillus mellis} and \textit{Gilliamella apicola}. 
    % \textit{Apilactobacillus} and \textit{Bombilactobacillus} are genera separated from \textit{Lactobacillus} recently \citep{zheng2020taxonomic}.  
    % \textit{Bombiscardovia coagulans} is a member of Bifidobacteriaceae isolated from \textit{Bombus lapidarius} \citep{killer2010bombiscardovia}. 
    % \newline
    % Different sample types are different in abundance distribution of bacteria (Figure \ref{BacteriaHeatmap}). 
    % For honey bee samples, bacterial community in \textit{Bee\_Amellifera\_hv13\_2} is dominated by \textit{Fructobacillus fructosus} (approximate 70\% relative abundance), while the other two honey bee samples share a similar bacteria abundance distribution with almost no species with relative sequence abundance $> 20\%$. 
    % As for bumble bee samples, \textit{Bifidobacterium commune} is the most abundant bacterial species in all three samples (about $20\%$ in \textit{Bee\_Bimpatiens\_hv3\_1} and about $40\%$ in the other two samples). 
    % Two bumble bees from the same hive (\textit{Bee\_Bimpatiens\_hv4\_1} and 
    % \textit{Bee\_Bimpatiens\_hv4\_1}) share similarity in bacterial abundance distribution. 
    % In the flower eDNA sample, some bee-associated bacteria were identified, including \textit{Bartonella apis}, \textit{Bifidobacterium asteroides}, \textit{Bombilactobacillus mellis} and \textit{Gilliamella apicola}.
    \newline
    Composition of arthropods and plants indicating interactions within pollination networks. 
    Most arthropods identified are pollinators within \textit{Apis} and \textit{Bombus} (Figure \ref{ArthropodHeatmap}). 
    However, some of them might be considered as false positive. 
    For example, \textit{Apis cerana}, \textit{Apis dorsata} and \textit{Apis florea} are mainly found in Asia and unlikely to present in the area where samples were collected. 
    These might derive by similarity between genomes of \textit{Apis mellifera} and other \textit{Apis} species. 
    % It should be noted that host contamination was removed before profiling. 
    % As a result, in honey bee samples, \textit{Apis} species were rare and \textit{Bombus} species were abundant, while in bumble bee samples, \textit{Bombus} species were rare and \textit{Apis} species are abundant. 
    As for plant species (Figure \ref{PlantHeatmap}), they indicate foraging areas of bees. Several crop species were identified, including \textit{Brassica napus} (rape), \textit{Brassica oleracea}, \textit{Cicer arietinum} (chickpea), \textit{Glycine max} (soybean), \textit{Helianthus annuus} (sunflower), \textit{Nicotiana sylvestris} (flowering tabacco) and \textit{Raphanus sativus} (radish). 
    \newline
    Fungal and virus species were identified. 
    In fungal communities of most samples, \textit{Nosema ceranae}, a widespread bee pathogen, was the dominant species. (Figure \ref{FungusHeatmap}). 
    Besides, three yeast species (\textit{Clavispora lusitaniae}, \textit{Saprochaete ingens} and \textit{Wickerhamiella sorbophila}) were also found in bees.  
    % and one opportunistic pathogen (\textit{Malassezia restricta}). 
    As for viruses, most of them are phages or arthropod-infecting species (Figure \ref{ViruseHeatmap}). 
    Phage species include \textit{Bifidobacterium phage BitterVaud1} infecting bee-commensal bacterium \textit{Bifidobacterium asteroides} \citep{bonilla2020honey}; \textit{Bacteriophage sp.} infecting \textit{Pseudomonas aeruginosa} \citep{essoh2015investigation}, an opportunistic pathogen that might contaminate bees \citep{bailey1968honey,papadopoulou1992experimental,papadopoulou1993haemocyte}; and unclassified species within Myoviridae and Siphoviridae. 
    Listed arthropod-infecting viruses including \textit{Apis mellifera filamentous virus} and \textit{Bombus cryptarum densovirus} that infect bees; and several parvoviruses (\textit{Blattodean pefuambidensovirus 1}, \textit{Hemipteran scindoambidensovirus 1}, \textit{Hymenopteran scindoambidensovirus 1} and \textit{Orthopteran scindoambidensovirus 1}) \citep{penzes2020reorganizing}.
    % % family Myoviridae and Siphoviridae
    % Most of the others are arthropod-associated viruses. 
    % \textit{Apis mellifera filamentous virus} and \textit{Bombus cryptarum densovirus} are bee-infecting viruses \citep{clark1978filamentous,bailey1981properties,schoonvaere2018study}. 
    % Several arthropod-infecting parvoviruses were found, including \textit{Blattodean pefuambidensovirus 1}, \textit{Hemipteran scindoambidensovirus 1}, \textit{Hymenopteran scindoambidensovirus 1} and \textit{Orthopteran scindoambidensovirus 1} \citep{penzes2020reorganizing}.

      \begin{figure}[H]
        \centering
        \includegraphics[width=0.8\textwidth]{../Figures/RelativeAbundance_0_01_species_Bacteria.pdf}
        \caption{Heatmaps for bacterial species abundance distribution in all samples. 
        The relative abundance takes reads assigned to bacterial species as background. 
        Species with relative abundance smaller than 1\% in all samples are collapsed as "others".}
        \label{BacteriaHeatmap}
        \end{figure}

    The integrated pipeline also provides information on FGC content, which shows function potentiality of metagenome. 
    Here FGC content was represented by identified KOs, which were used for KEGG pathway inference in order to illustrate metabolic potentially of metagenomic samples. 
    Reads assigned to plants and arthropods were not involved in pathway inference since they are unlikely to represent living organisms. 
    The coverage of a pathway was calculated by the ratio between number of annotated KOs and total number of KOs involved in that pathway. 
    \newline
    Here concern is given to metabolism pathways of carbonhydrates and amino acids, which are crucial for bee health. 
    Inferred pathways indicate potential capability of metabolism of sugars including fructose, sucrose, mannose and galactose (Figure \ref{Carbohydrate}), and all ten essential amino acids for honey bees (\textit{i.e.} arginine, histidine, isoleucine, leucine, lysine, methionine, phenylalanine, threonine, tryptophan and valine) (Figure \ref{AminoAcid}) \citep{groot1953protein}.

    \subsection{Evaluation of performance of integrated pipeline in species identification}
    The performance of integrated pipeline in species identification was evaluated by comparing it with assembly-dependent method, using rarefied datasets. 
    In integrated pipeline (Figure \ref{Pipeline}), clean non-host reads are first assembled to into contigs and assigned to taxa. 
    % To address false negative results in assembly-dependent taxon search \citep{sharon2015accurate,vollmers2017comparing}, 
    Then a reference database composed of genomes of assembly-represented species is constructed. 
    Reads not aligned to the reference database are subjected to assembly-independent taxon search. 
    The reference database and assembly-independent search helped improve species identification in all three sample types, especially in simulations of low sequencing depth (Figure \ref{CompareSensitivity}).

      \begin{figure}[H]
        \begin{subfigure}
          \centering
          \textbf{a}
        \end{subfigure}
        \hspace{0.32\linewidth}
        \begin{subfigure}

          \textbf{b}
        \end{subfigure}
        \hspace{0.32\linewidth}
        \begin{subfigure}

          \textbf{c}
        \end{subfigure}
        \newline
        \hfill
        \begin{subfigure}
          \centering
          \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/ExtraSpecies_Honey.pdf}
        \end{subfigure}
        \hfill
        \begin{subfigure}
          \centering
          \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/ExtraSpecies_Bumble.pdf}
        \end{subfigure}
        \hfill
        \begin{subfigure}
          \centering
          \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/ExtraSpecies_Flower.pdf}
        \end{subfigure}
        \caption{Integrated pipeline improves the detection of species richness. 
        The horizontal axis represents sequencing depth, and the vertical axis represents growth rate of detected species richness comparing the integrated pipeline and assembly-dependent species identification. 
        Different sequencing depth was simulated by rarefaction.
        % the ratio between number of novel species detected by assembly-independent search and number of species detected by assembly-dependent search. 
        Sample type is shown in the top left of each subfigure.}
        \label{CompareSensitivity}
        \end{figure}

    \subsection{Optimal sequencing depth required for detection of species and function diversity}
    In order to determine the influence of sequencing depth and thus optimize sequencing depth for 
    analysing taxonomic and functional diversity, rarefaction analysis was conducted. 
    Different sequencing depth was simulated by randomly subsampling original datasets at proportions from 10\% to 90\% at an interval of 10\%. 
    The relationship between sequencing depth (clean read pair number) and species/KO diversity (Hill numbers of order 0, 1, 2) was quantified by fitting and averaging asymptotic species accumulation models. 
    The slope of the model reflects the increase rate of diversity. 
    The point at which it drops to a cut-off value provides an estimation of optimal sequencing depth. 
    Besides, the asymptote of the model as sequencing depth tends to infinity provides an estimation of total diversity. 
    \newline
    Rarefaction assumes that the original dataset provides an almost complete inventory, which can be verified by final slope of rarefaction curve for Hill number of order 0 (richness). 
    Figure \ref{RichnessAccumulation} shows rarefaction curves for species/KO richness and Table \ref{RichnessModelSlope} summarizes their final slopes. 
    For species diversity rarefaction, all bumble bee samples are sufficient, with final slopes $< 0.1$ and completeness (ratio between final richness and asymptote) $> 0.98$. 
    % Their rarefaction curves for species richness almost level off, with final slopes $< 0.1$ and completeness (ratio between final richness and asymptote) $> 98\%$. 
    As for honey bees, {Bee\_Amellifera\_hv15\_1} and \textit{Bee\_Amellifera\_wild\_1} are sufficient, while \textit{Bee\_Amellifera\_hv13\_2} is insufficient, with final slope $> 1$ and completeness $< 0.8$. 
    % As for species diversity rarefaction of honey bee samples, the final slopes of \textit{Bee\_Amellifera\_hv15\_1} and \textit{Bee\_Amellifera\_wild\_1} are smaller than 0.1 and their completeness are higher than 99\%, indicating almost all species present in them were detected. 
    % The final slope of \textit{Bee\_Amellifera\_hv13\_2} is 1.0958 and its completeness is 79.31\%, indicating insufficient sequencing. 
    For the flower eDNA sample, the final slope of species richness rarefaction curve is 10.8380 and its completeness is 1.32\%, indicating more sequencing effort is needed for species profiling. 
    As for function diversity rarefaction, the final slopes of all KO richness rarefaction curves are higher than 15, indicating no dataset can providing an almost complete inventory of KOs. 
    Thus, estimation of optimal sequencing depth was conducted for the combinations of two sample types (honey bee and bumble bee) and one study type (species profiling), based on five datasets (two honey bees and three bumble bees).
    \begin{figure}[H]
      \begin{subfigure}
        \centering
        \textbf{a}
      \end{subfigure}
      \hspace{0.32\linewidth}
      \begin{subfigure}
        \centering
        \textbf{b}
      \end{subfigure}
      \hspace{0.32\linewidth}
      \begin{subfigure}
        \centering
        \textbf{c}
      \end{subfigure}
      \newline
      \hfill
      \begin{subfigure}
        \centering
        \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_TaxonHill_0.pdf}
      \end{subfigure}
      \hfill
      \begin{subfigure}
        \centering
        \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_TaxonHill_0.pdf}
      \end{subfigure}
      \hfill
      \begin{subfigure}
        \centering
        \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_TaxonHill_0.pdf}
      \end{subfigure}
      \newline
      \begin{subfigure}
        \centering
        \textbf{d}
      \end{subfigure}
      \hspace{0.32\linewidth}
      \begin{subfigure}
        \centering
        \textbf{e}
      \end{subfigure}
      \hspace{0.32\linewidth}
      \begin{subfigure}
        \centering
        \textbf{f}
      \end{subfigure}
      \newline
      \hfill
      \begin{subfigure}
        \centering
        \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_FunctionHill_0.pdf}
      \end{subfigure}
      \hfill
      \begin{subfigure}
        \centering
        \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_FunctionHill_0.pdf}
      \end{subfigure}
      \hfill
      \begin{subfigure}
        \centering
        \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_FunctionHill_0.pdf}
      \end{subfigure}
      \caption{Rarefaction curves for species (a, b, c) or Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs) (d, e, f) richness (Hill number of order 0). 
      The horizontal axis represents sequencing depth, and the vertical axis represents richness. 
      Sample type is shown in the top left of each subfigure. 
      Note that the scale of horizontal axis in subfigure c and f is much smaller than that in other subfigures.}
      \label{RichnessAccumulation}
      \end{figure}

      \begin{table}[H]
        \centering
        \caption{Summary of final point of rarefaction curve for richness (Hill number of order 0) of species or Kyoto Encyclopedia of Genes and Genomes (KEGG) orthologies (KOs). 
      Type: indicates whether this row reports rarefaction curve for species or KO richness. 
      Depth: sequencing depth taking million read pairs as unit. 
      Observed richness: observed species/KO richness. 
      Expected richness: expected species/KO richness predicted by modelling rarefaction curve. 
      Final slope: final slope of rarefaction curve. 
      Asymptote: asymptote as sequencing depth tends to infinity, calculated by modelling rarefaction curve. 
      Completeness: ratio between expected richness and asymptote.}
      \label{RichnessModelSlope}
      \resizebox{500pt}{!}{%
      \begin{tabular}{cccccccc} %ccccccc
      \toprule
Sample                           &Type   &Depth &Observed richness  & Expected richness          & Final slope      & Asymptote    & Completeness \\ \midrule
\textit{Bee\_Amellifera\_hv13\_2}&species&35.28 &144 & 146.69      & 1.0958  & 184.94  & 0.7931 \\
\textit{Bee\_Amellifera\_hv15\_1}&species&45.05 &97  & 88.53       & 0.0344  & 88.84   & 0.9966 \\
\textit{Bee\_Amellifera\_wild\_1}&species&44.46 &179 & 175.89      & 0.0991  & 176.84  & 0.9946 \\
\textit{Bee\_Bimpatiens\_hv3\_1} &species&53.61 &96  & 89.00       & 0.0897  & 90.79   & 0.9803 \\
\textit{Bee\_Bimpatiens\_hv4\_1} &species&48.42 &63  & 62.06       & 0.0626  & 62.92   & 0.9862 \\
\textit{Bee\_Bimpatiens\_hv4\_2} &species&45.80 &87  & 85.59       & 0.0146  & 85.69   & 0.9988 \\
\textit{Flower\_eDNA}            &species&0.88  &43  & 40.04       & 10.8380 & 3034.88 & 0.0132 \\
\textit{Bee\_Amellifera\_hv13\_2}&KO     &35.28 &9289& 9284.45     & 38.6844 & 10641.27& 0.8725 \\
\textit{Bee\_Amellifera\_hv15\_1}&KO     &45.05 &9046& 9102.78     & 43.8717 & 10453.98& 0.8707 \\
\textit{Bee\_Amellifera\_wild\_1}&KO     &44.46 &7216& 7151.72     & 21.4368 & 14779.89& 0.4839 \\
\textit{Bee\_Bimpatiens\_hv3\_1} &KO     &53.61 &6872& 6769.80     & 21.1588 & 28111.79& 0.2408 \\
\textit{Bee\_Bimpatiens\_hv4\_1} &KO     &48.42 &5427& 5410.71     & 19.6932 & 7665.58 & 0.7058 \\
\textit{Bee\_Bimpatiens\_hv4\_2} &KO     &45.8  &7288& 7259.56     & 15.6309 & 7873.64 & 0.9220 \\
\textit{Flower\_eDNA}            &KO     &0.88  &3732& 3544.54     & 60.4367 & 3571.11 & 0.9926 \\
      \bottomrule
      \end{tabular}%
      }
      \end{table}

    % \begin{figure}[H]
    %   \begin{subfigure}
    %     \centering
    %     \textbf{a}
    %   \end{subfigure}
    %   \hspace{0.32\linewidth}
    %   \begin{subfigure}
    %     \centering
    %     \textbf{b}
    %   \end{subfigure}
    %   \hspace{0.32\linewidth}
    %   \begin{subfigure}
    %     \centering
    %     \textbf{c}
    %   \end{subfigure}
    %   \newline
    %   \hfill
    %   \begin{subfigure}
    %     \centering
    %     \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_TaxonHill_0.pdf}
    %   \end{subfigure}
    %   \hfill
    %   \begin{subfigure}
    %     \centering
    %     \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_TaxonHill_0.pdf}
    %   \end{subfigure}
    %   \hfill
    %   \begin{subfigure}
    %     \centering
    %     \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_TaxonHill_0.pdf}
    %   \end{subfigure}
      % \newline
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \textbf{d}
      % \end{subfigure}
      % \hspace{0.32\linewidth}
      % \begin{subfigure}
      %   \centering
      %   \textbf{e}
      % \end{subfigure}
      % \hspace{0.32\linewidth}
      % \begin{subfigure}
      %   \centering
      %   \textbf{f}
      % \end{subfigure}
      % \newline
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_TaxonHill_1.pdf}
      % \end{subfigure}
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_TaxonHill_1.pdf}
      % \end{subfigure}
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_TaxonHill_1.pdf}
      % \end{subfigure}
      % \newline
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \textbf{g}
      % \end{subfigure}
      % \hspace{0.32\linewidth}
      % \begin{subfigure}
      %   \centering
      %   \textbf{h}
      % \end{subfigure}
      % \hspace{0.32\linewidth}
      % \begin{subfigure}
      %   \centering
      %   \textbf{i}
      % \end{subfigure}
      % \newline
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_TaxonHill_2.pdf}
      % \end{subfigure}
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_TaxonHill_2.pdf}
      % \end{subfigure}
      % \hfill
      % \begin{subfigure}
      %   \centering
      %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_TaxonHill_2.pdf}
      % \end{subfigure}
      % \caption{Rarefaction curves for species richness (Hill number of order 0). 
      % The horizontal axis represents sequencing depth, and the vertical axis represents species richness. 
      % Sample type is shown in the top left of each subfigure.}
      % \label{SpeciesAccumulation}
      % \end{figure}

%     \begin{table}[H]
%       \centering
%       \caption{Summary of final point of rarefaction curve for species richness (Hill number of order 0). 
%       Depth: sequencing depth taking million read pairs as unit. 
%       OR: observed species richness. 
%       ER: expected species richness predicted by modeling rarefaction curve. 
%       FS: final slope of rarefaction curve. 
%       Asym: asymptote calculated by modeling rarefaction curve. 
%       Comp: completeness, represented by ratio between ER and Asym.}
%       \label{SpeciesRichnessModel}
%       \resizebox{\textwidth}{!}{%
%       \begin{tabular}{ccccccc} %ccccccc
%       \toprule
% Sample                   & Depth   & OR  & ER          & FS      & Asym    & Comp \\ \midrule
% \textit{Bee\_Amellifera\_hv13\_2} & 35.28   & 144 & 146.69      & 1.0958  & 184.94  & 0.7931 \\
% \textit{Bee\_Amellifera\_hv15\_1} & 45.05   & 97  & 88.53       & 0.0344  & 88.84   & 0.9966 \\
% \textit{Bee\_Amellifera\_wild\_1} & 44.46   & 179 & 175.89      & 0.0991  & 176.84  & 0.9946 \\
% \textit{Bee\_Bimpatiens\_hv3\_1}  & 53.61   & 96  & 89.00       & 0.0897  & 90.79   & 0.9803 \\
% \textit{Bee\_Bimpatiens\_hv4\_1}  & 48.42   & 63  & 62.06       & 0.0626  & 62.92   & 0.9862 \\
% \textit{Bee\_Bimpatiens\_hv4\_2}  & 45.80   & 87  & 85.59       & 0.0146  & 85.69   & 0.9988 \\
% \textit{Flower\_eDNA}             & 0.88    & 43  & 40.04       & 10.8380 & 3034.88 & 0.0132 \\
%       \bottomrule
%       \end{tabular}
%       }
%       \end{table}

      % \begin{figure}[H]
        % \begin{subfigure}
        %   \centering
        %   \textbf{a}
        % \end{subfigure}
        % \hspace{0.32\linewidth}
        % \begin{subfigure}
        %   \centering
        %   \textbf{b}
        % \end{subfigure}
        % \hspace{0.32\linewidth}
        % \begin{subfigure}
        %   \centering
        %   \textbf{c}
        % \end{subfigure}
        % \newline
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_FunctionHill_0.pdf}
        % \end{subfigure}
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_FunctionHill_0.pdf}
        % \end{subfigure}
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_FunctionHill_0.pdf}
        % \end{subfigure}
        % \newline
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \textbf{d}
        % \end{subfigure}
        % \hspace{0.32\linewidth}
        % \begin{subfigure}
        %   \centering
        %   \textbf{e}
        % \end{subfigure}
        % \hspace{0.32\linewidth}
        % \begin{subfigure}
        %   \centering
        %   \textbf{f}
        % \end{subfigure}
        % \newline
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_FunctionHill_1.pdf}
        % \end{subfigure}
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_FunctionHill_1.pdf}
        % \end{subfigure}
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_FunctionHill_1.pdf}
        % \end{subfigure}
        % \newline
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \textbf{g}
        % \end{subfigure}
        % \hspace{0.32\linewidth}
        % \begin{subfigure}
        %   \centering
        %   \textbf{h}
        % \end{subfigure}
        % \hspace{0.32\linewidth}
        % \begin{subfigure}
        %   \centering
        %   \textbf{i}
        % \end{subfigure}
        % \newline
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Honey_FunctionHill_2.pdf}
        % \end{subfigure}
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Bumble_FunctionHill_2.pdf}
        % \end{subfigure}
        % \hfill
        % \begin{subfigure}
        %   \centering
        %   \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Flower_FunctionHill_2.pdf}
        % \end{subfigure}
        % \caption{Rarefaction curves for KO richness (Hill number of order 0). 
        % The horizontal axis represents sequencing depth, and the vertical axis represents KO richness. 
        % Sample type is shown in the top left of each subfigure.}
        % \label{FunctionAccumulation}
        % \end{figure}

%       \begin{table}[H]
%         \centering
%         \caption{Summary of final point of rarefaction curve for KO richness (Hill number of order 0). 
%         Depth: sequencing depth taking million read pairs as unit. 
%         OR: observed KO richness. 
%         ER: expected KO richness predicted by modeling rarefaction curve. 
%         FS: final slope of rarefaction curve. 
%         Asym: asymptote calculated by modeling rarefaction curve. 
%         Comp: completeness, represented by ratio between ER and Asym.}
%         \label{FunctionRichnessModel}
%         \resizebox{\textwidth}{!}{%
%         \begin{tabular}{ccccccc}
%         \toprule
% Sample                           &Depth  & OR    & ER        & FS       & Asym    & Comp   \\ \midrule
% \textit{Flower\_eDNA}            &0.88   & 3732  & 3544.54   & 60.4367  & 3571.11 & 0.9926 \\
% \textit{Bee\_Bimpatiens\_hv3\_1} &53.61  & 6872  & 6769.80   & 21.1588  & 28111.79& 0.2408 \\
% \textit{Bee\_Bimpatiens\_hv4\_1} &48.42  & 5427  & 5410.71   & 19.6932  & 7665.58 & 0.7058 \\
% \textit{Bee\_Bimpatiens\_hv4\_2} &45.8   & 7288  & 7259.56   & 15.6309  & 7873.64 & 0.9220 \\
% % \textit{Bee\_Amellifera\_hv13\_1}&0.84   & 3341  & 3322.97   & 909.4389 & 6975.56 & 0.4764 \\
% \textit{Bee\_Amellifera\_hv13\_2}&35.28  & 9289  & 9284.45   & 38.6844  & 10641.27& 0.8725 \\
% \textit{Bee\_Amellifera\_hv15\_1}&45.05  & 9046  & 9102.78   & 43.8717  & 10453.98& 0.8707 \\
% \textit{Bee\_Amellifera\_wild\_1}&44.46  & 7216  & 7151.72   & 21.4368  & 14779.89& 0.4839 \\
%         \bottomrule
%         \end{tabular}%
%         }
%         \end{table}

    Optimal sequencing depth is estimated by the point at which the slope of rarefaction curve drops to a cut-off value. 
    Table \ref{MinDepth} summarizes estimations of optimal sequencing depth for detection of species diversity, using 0.5, 0.1, 0.05 and 0.01 as cut-off values.  
    When order $q$ of Hill number equals 0, \textit{i.e.} species abundances are not considered, cut-off value of 0.1 for slope of rarefaction curve is sufficient for providing completeness $> 97\%$ in all samples. 
    The average optimal sequencing depth are 40.33 million for honey bees and 42.49 million for bumble bees. 
    When species abundances are considered (order $q$ equals 1 or 2), cut-off value of 0.01 can provide completeness $> 95\%$ in most pairs of sample and $q$ value. 
    For honey bees, the average optimal sequencing depth are 18.57 million ($q = 1$) and 17.45 million ($q = 2$). 
    For bumble bee samples, the average optimal sequencing depth are 40.33 million ($q = 1$) and 24.77 million ($q = 2$).

    \begin{table}[H]
      \centering
      \caption{Summary of optimal sequencing depth estimated from different cut-off values of slope. 
      Optimal sequencing depth was estimated by the point at which the slope of rarefaction curve drops to a cut-off value (0.5, 0.1, 0.05 or 0.01). 
      $q$: order of Hill number that determines sensitivity to species abundances. 
      Asymptote: asymptote as sequencing depth tends to infinity, calculated by modelling rarefaction curve. 
      Depth: optimal sequencing depth taking million read pairs as unit.  
      Expected diversity: expected Hill number provided by the optimal sequencing depth.   
      Completeness: ratio between expected diversity and asymptote.}
      \label{MinDepth}
      \resizebox{500pt}{!}{%
      \begin{tabular}{ccccccccccccccc}
      \hline
      \multirow{2}{*}{Sample} & \multirow{2}{*}{\textit{q}} & \multirow{2}{*}{Asymptote} & \multicolumn{3}{c}{Slope $ < 0.5$} & \multicolumn{3}{c}{Slope $< 0.1$} & \multicolumn{3}{c}{Slope $< 0.05$}  & \multicolumn{3}{c}{Slope $< 0.01$}  \\ \cline{4-15} & & & Depth & \makecell[c]{Expected\\diversity} &Completeness & Depth & \makecell[c]{Expected\\diversity} & Completeness & Depth & \makecell[c]{Expected\\diversity} & Completeness & Depth & \makecell[c]{Expected\\diversity} & Completeness \\ \hline
      \textit{Bee\_Amellifera\_hv15\_1}&0&88.84  &23.54&84.85 &0.9551  &36.27&87.99 &0.9905  &41.95&88.40 &0.9951  &55.75&88.74 &0.9989\\
      \textit{Bee\_Amellifera\_wild\_1}&0&176.84 &30.53&172.46&0.9752  &44.39&175.88&0.9946  &50.57&176.33&0.9971  &65.86&176.70&0.9992\\
      \textit{Bee\_Amellifera\_hv15\_1}&1&12.28  &3.18 &10.67 &0.8699  &5.83 &11.33 &0.9224  &7.21 &11.43 &0.9304  &15.21&11.58 &0.9429\\
      \textit{Bee\_Amellifera\_wild\_1}&1&14.77  &0.69 &13.49 &0.9137  &3.30 &14.03 &0.9498  &6.20 &14.23 &0.9635  &21.93&14.58 &0.9868\\  
      \textit{Bee\_Bimpatiens\_hv3\_1}&0&90.79  &30.80&83.69&0.9218  &52.03&88.85&0.9786  &62.56&89.61&0.9870  &92.35&90.32&0.9948\\
      \textit{Bee\_Bimpatiens\_hv4\_1}&0&62.92  &24.83&57.22&0.9093  &42.76&61.61&0.9791  &51.24&62.22&0.9887  &73.55&62.76&0.9973\\
      \textit{Bee\_Bimpatiens\_hv4\_2}&0&85.69  &21.75&82.30&0.9604  &32.67&85.01&0.9920  &37.39&85.35&0.9960  &48.40&85.62&0.9992\\

      \textit{Bee\_Bimpatiens\_hv3\_1}&1&21.59  &3.63 &10.50&0.4862  &9.47 &11.76&0.5449  &15.37&12.18&0.5639  &58.09&13.04&0.6040\\
      \textit{Bee\_Bimpatiens\_hv4\_1}&1&11.85  &3.80 &9.45 &0.7973  &10.15&10.87&0.9174  &14.68&11.19&0.9445  &31.22&11.57&0.9766\\
      \textit{Bee\_Bimpatiens\_hv4\_2}&1&12.66  &4.17 &10.24&0.8087  &10.47&11.67&0.9213  &14.85&11.98&0.9458  &31.68&12.36&0.9760\\

      \textit{Bee\_Bimpatiens\_hv3\_1}&2&8.73   &2.20 &6.16 &0.7051  &6.28 &7.06 &0.8088  &9.45 &7.28 &0.8344  &23.89&7.60 &0.8708\\
      \textit{Bee\_Bimpatiens\_hv4\_1}&2&8.33   &3.12 &6.28 &0.7540  &8.52 &7.50 &0.9002  &12.30&7.76 &0.9323  &26.68&8.09 &0.9718\\
      \textit{Bee\_Bimpatiens\_hv4\_2}&2&7.12   &2.90 &5.34 &0.7493  &7.47 &6.38 &0.8961  &10.58&6.60 &0.9270  &23.73&6.90 &0.9683\\          
      \hline
      \end{tabular}%
      }
      \end{table}

%       \begin{table}[H]
%         \centering
%         \caption{Summary of minimal sequencing depth estimated from different cut-off values of slope. 
%         Minimal sequencing depth was estimated by the point at which the slope of rarefaction curve drops to a cut-off value (0.5, 0.1, 0.05 or 0.01), which is indicated in names of columns. 
%         For example, MinD\_0.5 represents estimated minimal sequencing depth taking 0.5 as cut-off value, and ED\_0.5 is expected Hill number from MinD\_0.5. Comp\_0.5 is the ratio between OptDiv\_0.5 and the asymptote (represented by Asym). 
%         Unit of sequencing depth is million read pairs. 
%         $q$ refers to order of Hill number determining sensitivity to species abundance distribution.}
%         \label{MinDepth}
%         \resizebox{500pt}{!}{%
%         \begin{tabular}{ccccccccccccccc} % lllllllllllllll
%         \toprule
% Sample&$q$&Asym  &MinD\_0.5&ED\_0.5&Comp\_0.5  &MinD\_0.1&ED\_0.1&Comp\_0.1  &MinD\_0.05&ED\_0.05&Comp\_0.05  &MinD\_0.01&ED\_0.01&Comp\_0.01 \\ \midrule
% % \textit{Bee\_Amellifera\_hv13\_1}&0&4262.21 &1373.18&4022.37&0.9437 &2142.41&4213.53&0.9886 &2473.90&4237.44&0.9942 &3244.93&4256.60&0.9987\\
% % \textit{Bee\_Amellifera\_hv13\_2}&0&184.94  &59.87&165.10&0.8927  &120.88&179.94&0.9730  &149.83&182.02&0.9842&226.97&183.88&0.9942\\
% \textit{Bee\_Amellifera\_hv15\_1}&0&88.84  &23.54&84.85 &0.9551  &36.27&87.99 &0.9905  &41.95&88.40 &0.9951  &55.75&88.74 &0.9989\\
% \textit{Bee\_Amellifera\_wild\_1}&0&176.84 &30.53&172.46&0.9752  &44.39&175.88&0.9946  &50.57&176.33&0.9971  &65.86&176.70&0.9992\\

% % \textit{Bee\_Amellifera\_hv13\_1}&1&7.06  &0.06&6.56&0.9296  &0.63&6.66&0.9434  &1.30&6.71&0.9501  &3.41&6.76&0.9571\\
% % \textit{Bee\_Amellifera\_hv13\_2}&1&16.01 &5.55&15.03&0.9385 &8.70&15.81&0.9872 &10.08&15.91&0.9934  &13.34&15.99&0.9984\\
% \textit{Bee\_Amellifera\_hv15\_1}&1&12.28  &3.18 &10.67 &0.8699  &5.83 &11.33 &0.9224  &7.21 &11.43 &0.9304  &15.21&11.58 &0.9429\\
% \textit{Bee\_Amellifera\_wild\_1}&1&14.77  &0.69 &13.49 &0.9137  &3.30 &14.03 &0.9498  &6.20 &14.23 &0.9635  &21.93&14.58 &0.9868\\

% % \textit{Bee\_Amellifera\_hv13\_1}&2&4.85   &0.02 &4.26  &0.8789  &0.20 &4.29 &0.8848  &0.94&4.34 &0.8950  &4.27 &4.42 &0.9115\\
% % \textit{Bee\_Amellifera\_hv13\_2}&2&11.00  &4.47 &10.15 &0.9232  &7.20 &10.83&0.9845  &8.38&10.91&0.9922  &11.14&10.98&0.9984\\
% \textit{Bee\_Amellifera\_hv15\_1}&2&8.40   &3.10 &6.35  &0.7561  &8.17 &7.49  &0.8909  &11.97&7.75  &0.9229  &27.99&8.11 &0.9655\\
% \textit{Bee\_Amellifera\_wild\_1}&2&9.88   &0.18 &9.49  &0.9602  &0.90 &9.63  &0.9747  &1.73 &9.69  &0.9805  &6.91 &9.80  &0.9916\\

% \textit{Bee\_Bimpatiens\_hv3\_1}&0&90.79  &30.80&83.69&0.9218  &52.03&88.85&0.9786  &62.56&89.61&0.9870  &92.35&90.32&0.9948\\
% \textit{Bee\_Bimpatiens\_hv4\_1}&0&62.92  &24.83&57.22&0.9093  &42.76&61.61&0.9791  &51.24&62.22&0.9887  &73.55&62.76&0.9973\\
% \textit{Bee\_Bimpatiens\_hv4\_2}&0&85.69  &21.75&82.30&0.9604  &32.67&85.01&0.9920  &37.39&85.35&0.9960  &48.40&85.62&0.9992\\

% \textit{Bee\_Bimpatiens\_hv3\_1}&1&21.59  &3.63 &10.50&0.4862  &9.47 &11.76&0.5449  &15.37&12.18&0.5639  &58.09&13.04&0.6040\\
% \textit{Bee\_Bimpatiens\_hv4\_1}&1&11.85  &3.80 &9.45 &0.7973  &10.15&10.87&0.9174  &14.68&11.19&0.9445  &31.22&11.57&0.9766\\
% \textit{Bee\_Bimpatiens\_hv4\_2}&1&12.66  &4.17 &10.24&0.8087  &10.47&11.67&0.9213  &14.85&11.98&0.9458  &31.68&12.36&0.9760\\

% \textit{Bee\_Bimpatiens\_hv3\_1}&2&8.73   &2.20 &6.16 &0.7051  &6.28 &7.06 &0.8088  &9.45 &7.28 &0.8344  &23.89&7.60 &0.8708\\
% \textit{Bee\_Bimpatiens\_hv4\_1}&2&8.33   &3.12 &6.28 &0.7540  &8.52 &7.50 &0.9002  &12.30&7.76 &0.9323  &26.68&8.09 &0.9718\\
% \textit{Bee\_Bimpatiens\_hv4\_2}&2&7.12   &2.90 &5.34 &0.7493  &7.47 &6.38 &0.8961  &10.58&6.60 &0.9270  &23.73&6.90 &0.9683\\

% % \textit{Flower\_eDNA}&0&3034.88 &922.36&2870.43&0.9458  &1428.16&2996.00&0.9872  &1648.07&3011.86&0.9924  &2173.39&3024.82& 0.9967\\
% % \textit{Flower\_eDNA}&1&6.46    &0.74  &5.59   &0.8656  &2.89   &6.07   &0.9399  &4.50   &6.19   &0.9575  &11.30  &6.34   & 0.9810\\
% % \textit{Flower\_eDNA}&2&4.40    &0.55  &3.50   &0.7940  &2.47   &3.92   &0.8910  &3.93   &4.03   &0.9145  &10.24  &4.17   & 0.9463\\
%         \bottomrule
%         \end{tabular}%
%         }
%         \end{table}

\section{Discussion}
% Shotgun metagenomics provides powerful tools for microbiome investigations, yet its utilization is hindered by bioinformatical challenges and high cost of deep sequencing. 
Here, I constructed an integrated pipeline combining assembly-dependent and -independent methods for taxonomic and functional profiling of shotgun metagenomics, and applied it to analysis of metagenomes from honey bees, bumble bees and flower washes. 
The profiling results showed that the integrated pipeline is able to determine a comprehensive view of the taxonomic diversity (across multiple clades) and infer functional potentiality of metagenome by providing information on FGC content. 
The developed method is more sensitive to species identification compared with standard assembly-dependent methods. 
Then I computed expected species/FGC diversity represented by given sequencing depth through rarefaction analysis, in order to optimize sequencing depth to balance cost and reliability of analysis results. 
Optimal sequencing depth differs by types of samples and investigations. 
For estimation of species diversity, sequencing depth can be optimized especially when low leverage is given to rare species. 
As for functional profiling, deeper sequencing depth is required.  
% Then I simulated the relationship between species/KO diversity measured by Hill numbers (order $q = 0, 1, 2$) and sequencing depth measured by clean read pair ($2 \times 150$ bp) number. 
% It is showed that to detect species diversity, the average minimal sequencing depth for honey bee samples are 40.33 million ($q = 0$), 18.57 million ($q = 1$) and 17.45 million ($q = 2$); for bumble bee samples, the averages are 42.49 million ($q = 0$), 40.33 million ($q = 1$) and 24.77 million ($q = 2$); and for flower surface, more than 0.88 million sequencing depth is needed. 
% As for function diversity, no datasets in this project is sufficient for providing an almost complete inventory of KOs, indicating deeper sequencing is needed, \textit{i.e.} sequencing depth need to be higher than 45.06 million for honey bees, 53.61 million for bumble bees and 0.88 million for flower surface. 
% Additionally, bee symbionts and pathogens were found on the flower eDNA sample, and pathway inference indicates that bee-associated microbiome may have capability of metabolizing amino acids that are essential for host.
\newline
The integrated pipeline for taxonomic and functional profiling of metagenome provides several advantages. 
First, through combination with assembly-free taxon search, it helps solve high false negative rate associated with assembly-dependent species identification, which is caused by reads left unassembled. 
This was shown by analysing real metagenomic datasets from pollination system and simulating different sequencing depth by rarefaction. 
Second, the modularity of integrated pipeline provides flexibility for incorporation of alternative tools. 
For example, BWA aligner \citep{li2009fast} serves as an alternative for Bowtie2, and SPAdes can be replaced by other metagenomic assemblers such as Megahit \citep{li2015megahit} and IDBA-UD \citep{peng2012idba}. 
Third, the output files generated by each step are recorded and can be inspected easily, which provides transparency for troubleshooting. 
However, the performance of integrated pipeline would to be evaluated more comprehensively by further analysis of different datasets and comparison with other profiling strategies. 
Metagenomes from other host species could be used to illustrate the performance of integrated pipeline. 
Standardized mock metagenomes, even though they are less complex than real ones, could also be used to benchmark sensitivity and accuracy of integrated pipeline since it is an artificial metagenome with predefined diversity, generated by combining known sequences from different species \citep{vollmers2017comparing,sczyrba2017critical,becker2020modular}. 
Further, the performance of integrated pipeline in metagenome profiling could be further benchmarked against other analysis strategies, such as MG-RAST \citep{meyer2008metagenomics}, SqueezeMeta \citep{tamames2019squeezemeta} and Kraken \citep{wood2014kraken}. 
% High throughput sequencing provides approaches including amplicon sequencing and shotgun metagenomics for investigations of bee-associated microbiome, which imposes significant impact on bee health \citep{engel2016bee,raymann2018role,bonilla2018functional}. 
% a crucial issue receiving increasing concern \citep{breeze2011pollination,hristov2020significance}. 
% Shotgun metagenomics provides unique advantages over amplicon sequencing, which is vastly used in bee microbiome investigations (\textit{e.g.} \cite{geldert2021dietary,wang2021gut,powell9field,kapheim2021composition}). 
% In amplicon sequencing, a species-specific barcode region is amplified and sequenced \citep{abdelfattah2018metabarcoding}. 
% Since it captures a small region of the whole genome, amplicon sequencing dataset is relatively small in size, which makes it low-cost both economically and computationally. 
% Besides, comprehensive and streamlined softwares for amplicon sequencing analysis are available, including QIIME \citep{bolyen2019reproducible}, Mothur \citep{schloss2020reintroducing} and VSEARCH \citep{rognes2016vsearch}. 
% However, a single amplicon sequencing dataset does not represent all taxa in the community since different barcode region is used for identification of different taxonomic group, \textit{e.g.} 
% 16S ribosomal RNA (rRNA) for bacteria \citep{hayashi2002phylogenetic,eckburg2005diversity}, 
% internal transcribed spacer (ITS) for fungi \citep{nilsson2008intraspecific}, 
% cytochrome c oxidase subunit I (COI) for Animalia \citep{hebert2003barcoding} 
% and plastid genes for plants \citep{group2009dna}. 
% Besides, it does not provide information on FGCs. 
% Metabolic capacity of microbiome need to be inferred based on reference genomes \Citep{asshauer2015tax4fun,douglas2018predicting}. 
% Shotgun metagenomics provides an alternative to overcome these drawbacks by capturing and sequencing DNA in a sample unselectively. 
% It is capable of representing all taxonomic groups and providing information on FGC composition. 
% \newline
% The advantages of shotgun metagenomics over amplicon sequencing are valuable for investigations in bee-associated microbiome for two reasons. 
% First, bees visit numerous niches during environment exploration and foraging activities, and can get contact with diverse eDNA signatures, which provide insight into bee ecology by reflecting interactions between bees and other organisms including plants, arthropods, bacteria and viruses \citep{bovo2018shotgun,ribani2020honey,bovo2020shotgun,matsuzawa2020application}. 
% These materials represent a diverse assemblage which is difficult to be profiled comprehensively via amplicon sequencing. 
% Besides, bee bacterial symbionts are diversified at strain level \citep{engel2012functional,powell2016strain,ellegaard2020vast}. 
% Bacterial strains are often highly variable in gene content \citep{cordero2014explaining,brockhurst2019ecology}, and thus different in metabolic capacity. 
% However, resolving strain-level diversity by species-specific region is difficult \citep{rodriguez2018much,ciufo2018using,olm2020consistent}, and no information on gene content is provided by amplicon sequencing. 
% Based on these considerations, shotgun metagenomics provides valuable tools for bee microbiome investigations. 
% However, utilization of shotgun metagenomics is hindered by unique challenges. 
% First, there is not a golden standard for bioinformatics of metagenomic annotation pipeline to provide abundances of taxa and FGCs. 
% Second, shotgun metagenomics is often associated with high cost since it captures a big proportion of microbial genomes. 
\newline
The integrated pipeline was used to analyse metagenomes of bees and flower washes, illustrating their taxon composition and metabolic potentiality. 
Bee-associated microorganisms were identified in flower washes, including both pathogens (\textit{e.g.} \textit{Nosema ceranae} and \textit{Apis mellifera filamentous virus}) and symbionts (\textit{e.g.} \textit{Bifidobacterium asteroides}, \textit{Bombilactobacillus mellis} and \textit{Gilliamella apicola}). 
Flowers have been indicated as hubs for transmission of bee pathogens \cite{durrer1994shared,koch2017role,graystock2020dominant}. 
Through shared flower use, multiple pathogens transmit between pollinators, including \textit{Nosema ceranae} \citep{graystock2015parasites} 
% furst2014disease
and \textit{Crithidia bombi} \citep{figueroa2019bee}. 
There is also growing evidence for present of bee symbionts on flower \citep{mcfrederick2012environment,mcfrederick2017flowers,keller2020more,vannette2020floral}.
However, its role in bee microbiome assembly and functional importance in pollination remains an open question \citep{keller2020more,vannette2020floral}. 
As for function profiling, it is indicated that bee-associated microbiome is capable of metabolizing carbohydrates such as glucose, fructose, sucrose and mannose. 
Carbohydrates are main component of bee diet and microbiome-mediated carbohydrate-processing have been vastly investigated \citep{engel2012functional,lee2015saccharide,lee2018differential,taylor2019effect}. 
Besides, pathways for metabolism of all ten essential amino acids for honey bees \citep{groot1953protein} were inferred. 
Essential amino acids are crucial for bee health \citep{simcock2014single,paoli2014dietary,stabler2015nutrient,hendriksma2019effects} and influence feeding preference due to their potential deficiency in single pollen source \citep{cook2003honey,hendriksma2014amino,hendriksma2016honey}. 
Whether bee-associated microbiome influences host health by providing amino acids need to be further investigated. 
% deficiency of arginine in dandelion \citep{herbert1970brood} and histidine in corn \citep{hocherl2012evaluation}.
\newline
Optimization of sequencing depth is important for shotgun metagenomics since insufficient sequencing causes underestimation of taxonomic/functional diversity \citep{cattonaro2018you,zaheer2018impact,gweon2019impact,pereira2019impact}, while deep sequencing is of high cost. 
Here, expected species/KO diversity provided by given sequencing depth was estimated by rarefaction and model fitting, using datasets from honey bees, bumble bees and flower washes. 
It was shown that increasing sequencing depth boosted identification of species/KO, highlighting value of deep metagenomic sequencing. 
For function profiling, no dataset involved here is big enough to provide an almost complete inventory of KOs even though most samples were deeply sequenced, indicating such task is demanding about sequencing depth. 
Therefore, when dealing with function potentiality of microbiome associated with bees or flowers, retrieving as many reads as possible would be recommended. 
As for species profiling, although deep sequencing is still valuable, sequencing depth can be optimized when the budget is limited, especially when low leverage is given to rare species. 
For honey bees, approximate 40 million sequencing depth (12 Gbp) can be sufficient for representing species richness. 
When assessing biodiversity with reduced emphasis on rare species, 17-19 million sequencing depth (5.1-5.7 Gbp) can provide robust estimation. 
As for bumble bees, about 40-43 million sequencing depth (12-12.9 Gbp) can provide reliable estimation for species richness and diversity index with emphasis on species of general abundance, and about 25 (7.5 Gbp) million sequencing depth can be sufficient for estimation of biodiversity with high leverage on abundant species. 
It should be noted that the efficacy of optimal sequencing depth estimation is reduced by limited number of samples. 
Here, honey bee and bumble bee associated microbiome was represented by only three samples, and there was only one sample of flower eDNA. 
However, microbiomes of pollination system are highly dynamic and variable in diversity. 
To generate reliable guideline for sequencing depth optimization, more samples need to be evolved. 
Besides, there is a lack of repeat in sequencing depth subsampling. 
Repeated subsampling boosts precise computation of rarefaction curve and increases accuracy of estimation of optimal sequencing depth.

\section{Conclusion}
Shotgun metagenomics is capable of illustrating diversity of multiple taxonomic clades and gene content, and thus provides unique advantages over vastly used amplicon sequencing, particularly for investigations of highly diverse microbial communities. 
However, utilization of shotgun metagenomics is hindered by challenges in data analysis and high cost of sequencing. 
Here, I constructed an integrated pipeline for analysis of shotgun metagenomic data. 
It provides benefits in terms of results, flexibility and transparency. 
% The pipeline was used for profiling metagenomic datasets from pollination system. 
I also constructed a framework for optimizing depth of shotgun metagenomic sequencing in order to balance high cost of sequencing and reliability of analysis results. 
The pipeline and the framework were used for analysis of real datasets from pollination system. 
For species diversity detection, about 40 million clean $2 \times 150$ bp read pairs would be sufficient honey bee samples to balance sequencing cost and reliable output; and 43 million would be sufficient for sequencing bumble bee samples. 
Shallower sequencing can be adopted with reduced emphasis on rare species. 
As for functional profiling, obtaining as many reads as possible would be the recommendation. 
These results provide guidelines for cost-effective metagenomic investigations of bee microbiomes. 
Methods used in this project can be adopted to similar studies for other host species, which are recommended before undertaking metagenomic projects with big sample size.

% Then I conducted rarefaction analysis, computing expected species/KO diversity (Hill numbers of order 0, 1, 2) returned by given sequencing effort in order to optimize sequencing depth for balancing cost and reliable estimation of diversity. 
% It was shown that functional profiling is demanding about sequencing effort and obtaining as many sequencing reads as possible would be recommended. 
% For taxonomic profiling, deep sequencing is still valuable but sequencing depth can be optimized, especially when little emphasis is given to rare taxa. 
% Sequencing depth of 45.06 million (13.52 Gbp) for honey bees, 53.61 million (16.08 Gbp) for bumble bees and 0.88 million (0.26 Gbp) for flower surface were not sufficient for providing almost complete inventories of KOs. 
% For species diversity of order 0, 1 and 2, the average minimal sequencing depth for honey bee samples are 40.33 million (12.10 Gbp), 18.57 million (5.57 Gbp) and 17.45 million (5.24 Gbp), respectively; for bumble bee samples, the averages are 42.49 million (12.75 Gbp), 40.33 million (12.10 Gbp) and 24.77 (7.43 Gbp) million. 
% As for species profiling of flower surface, more than 0.88 million (0.26 Gbp) sequencing depth is needed. 
% These results are helpful for applying shotgun metagenomics to pollination system investigations in an cost-effective way. 
% The integrated pipeline and the framework for rarefaction analysis can be used for researches on other environments.

% bee importance {gallai2009economic,gill2016protecting,ferrier2018economic,cooley2021buzz}, 
% bee decline {cappa2021analysis,frizzera2021interaction,hemberger2021historical}

\section{Data and Code Availability}
Scripts used for the analyses are available at \href{https://github.com/CongLiu37/CMEECourseWork/tree/main/MResProject}{github}. 
Metagenomic datasets involved in this project cannot be made available publicly since they are yet to be formally published.

\newpage

  % \bibliographystyle{unsrt}
  % \bibliographystyle{plainnat}
  \bibliography{reference}

\newpage

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
\section{Supplementary}
  \beginsupplement
  \subsection{Parameter settings of integrated pipeline}
  \label{ParameterSetting}
    \subsubsection{Quality control and host removing}
    \label{HostRemove}
    Raw data quality control was conducted to reduce compromise from low-quality reads.
    Raw data quality was checked using FastQC v.0.11.5 \citep{andrews2010fastqc} before filtering. 
    FastQC reports of raw reads showed the following aspects need to be covered in quality control: (1) low base quality in 3'-end (Figure \ref{QC}a); (2) uneven base content in 5'-end (Figure \ref{QC}c) and (3) the present of Nextera adaptors (Figure \ref{QC}e). 
    Therefore, raw reads were filtered using Trimmomatic v.0.39 \citep{bolger2014trimmomatic}, which (1) trimmed adaptors; (2) cutted 15 bases from the 5'-ends of reads; (3) cut bases off from 3'-ends of reads if Phred-33 quality is below 20; (4) dropped reads shorter than 50 bp; (5) dropped reads if average Phred-33 quality is below 20. 
    Then unpaired reads were removed and quality of clean data was checked using FastQC (Figure \ref{QC}d, \ref{QC}e, \ref{QC}f).
    \newline  
    After quality control, host read pairs were removed. 
    First, clean read pairs were mapped to host genome (GCA\_003254395.2 for \textit{Apis mellifera} and GCA\_000188095 for \textit{Bombus impatiens}, downloaded from NCBI) using Bowtie2 v.2.4.2 \citep{langmead2012fast} with flags \textit{--end-to-end} and \textit{--sensitive}. 
    With flag \textit{--end-to-end}, Bowtie2 requires the read aligned without any clipping from neither end, and \textit{--sensitive} maintains a trade-off between speed and sensitivity. 
    SAM files generated by Bowtie2 were converted to BAM format using SAMtools v.1.11 \citep{li2009sequence}. 
    Then non-host read pairs were extracted from BAM files by SAMtools according to the present of SAM flag 12 (neither forward nor reverse read in a pair of reads is mapped).

    \subsubsection{Assembly-dependent species identification}
    \label{AssemblyDependent}
    In order to identify taxa of metagenome, \textit{de novo} assembly was conducted. 
    Non-host read pairs were assembled using SPAdes v.3.15.2 \citep{prjibelski2020using}. 
    Values of k-mer ranged from 21 to 101 at interval of 10. 
    Flags \textit{--only-assembler} and \textit{--meta} were used. 
    Through flag \textit{--meta}, SPAdes runs metaSPAdes which is developed for metagenomic assembly \citep{nurk2017metaspades}. 
    The \textit{--only-assembler} flag skips read error correction and runs assembly only. 
    Its utilization is justified by the following facts. 
    First, when \textit{--only-assembler} is not used, SPAdes conducts error correction before assembly. 
    It is conducted by BayesHamme, which is optimized for single cell sequencing instead of shotgun metagenomics \citep{nikolenko2013bayeshammer}. 
    Besides, reads used for \textit{de novo} assembly had been filtered to ensure their quality.
    \newline
    After assembly, taxon identification was conducted using DIAMOND v.2.0.7.145 \citep{buchfink2015fast} and MEGAN6 \citep{huson2007megan}. 
    Assembled contigs with a length above 500 bp were aligned to nr database using DIAMOND v.2.0.7.145 with \textit{--long-reads} flag. 
    This flag triggers frame-shift aware alignment mode, which is optimized for long sequence alignment. 
    Therefore, short contigs (length $<$ 500 bp) were not retained. 
    Besides, alignments with an E-value $< 1e-5$ or identity $< 50\%$ were removed, and for each contig, only alignments above 10\% of the best local bit score were retained. 
    The output of DIAMOND was analysed by the \textit{blast2rma} tool of MEGAN6 with \textit{-lg} flag, which runs lowest-common-ancestor (LCA)-based algorithm developed for long contigs and assigns each contig to a taxon \citep{huson2018megan}. 
    The parameter \textit{-supp} was 0, which means the present of a taxon would be identified as long as at least one contig was assigned. 
    This value was used because a contig is assembled from multiple short reads and represents a strong signal for the present of a taxon.

    \subsubsection{Fragment recruitment}
    To integrate individual genomic data of species identified by assembly-dependent search, a reference database comprising reference genome dataset, \textit{i.e.} genomic sequences in FASTA format and corresponding gff file, was constructed. 
    For each species represented by assembly, its reference genome dataset, if available, was downloaded from NCBI using its \textit{datasets} command-line tool and added to the reference database. 
    \newline
    Then fragment recruitment was conducted. 
    The non-host read pairs were mapped to genomic sequences in the reference database using Bowtie2. 
    Read pairs that were not recruited were extracted using SAMtools. 
    Settings for Bowtie2 and SAMtools were the same as that described in \ref{HostRemove}. 
    Read pairs recruited by the reference database were assigned to corresponding species, while the others were subjected to assembly-independent search.

    \subsubsection{Assembly-independent species identification}
    In order to detect species not represented by assembly \citep{sharon2015accurate,vollmers2017comparing}, assembly-independent search was conducted, taking read pairs not recruited by the reference database as input. 
    These reads were aligned to nr database through DIAMOND without using \textit{--long-reads} flag, which triggers computing alignments for short metagenomic reads. 
    Other settings were the same as described in \ref{AssemblyDependent}. 
    Then the output of DIAMOND was analysed by MEGAN6 (\textit{blast2rma} tool), which assigns read pairs to taxa through LCA algorithm. 
    Here the parameter \textit{-supp} was 0.1, which means a taxon is reported after being represented by at least 0.1\% of all assigned read pairs. 
    It was used in order to avoid false positive results.

    \subsubsection{Functional annotation}
    Functional annotation was conducted by EggNOG-mapper v.2.1.2 \citep{huerta2017fast}. 
    Sequences were searched against eggNOG database \citep{huerta2019eggnog} for best seed orthologs using DIAMOND and fine-grained orthology assignments were retrieved from pre-computed eggNOG phylogenetic trees. 
    Then functional descriptions of retrieved orthologs including Gene Ontology (GO) terms \citep{gene2004gene}, KOs, Enzyme Commission (EC) numbers \citep{webb1992enzyme}, Carbohydrate-Active Enzymes (CAZy) terms \citep{cantarel2009carbohydrate} and Clusters of Orthologous Groups (COG) functional categories \citep{tatusov2000cog} were transferred to query sequences.

  \subsection{Candidate models for fitting rarefaction curves}
  \begin{table}[H]
    \centering
    \caption{Candidate species accumulation models. 
    Dependent variable $D^{(q)}$ is Hill number of order $q$ and 
    independent variable $x$ is sequencing depth. 
    $a$, $b$, $c$, $d$ are fitted coefficients.}
    \label{candidate_models}
    \resizebox{500pt}{!}{%
    \begin{tabular}{lllllll}
    \toprule
    Model & Parameter($k$) & Derivative & Asymptote & Reference \\ \midrule
    $D^{(q)} = \frac{ax}{bx+1}$ & 2 & $\frac{dD^{q}}{dx} = \frac{a}{(bx+1)^2}$ & $\frac{a}{b}$ & \cite{clench1979make}  \\
    $D^{(q)} = a(1-e^{-bx})$ & 2 & $\frac{dD^{q}}{dx} = abe^{-bx}$ & $a$ & \cite{miller1989documenting}  \\
    $D^{(q)} = a-bc^x$ & 3 & $\frac{dD^{q}}{dx} = -bc^{x}log(c)$ & $a$ & \cite{ratkowsky1983nonlinear}  \\
    $D^{(q)} = a(1-e^{-bx})^c$ & 3 & $\frac{dD^{q}}{dx} = abce^{-bx}(1-e^{-bx})^{c-1}$ & $a$ & \cite{ratkowsky1990handbook}  \\
    $D^{(q)} = a(1-(1+(\frac{x}{c})^d)^{-b})$ & 4 & $\frac{dD^{q}}{dx} = \frac{abd}{c}(\frac{x}{c})^{d-1}(1+(\frac{x}{c})^d)^{-b-1}$ & $a$ & \cite{mielke1974some}\\
    \bottomrule
    \end{tabular}
    }
    \end{table}

  \subsection{Exemplification of effect of quality control}
  \begin{figure}[H]
    \begin{subfigure}
      \centering
      \textbf{a}
    \end{subfigure}
    \hspace{0.32\linewidth}
    \begin{subfigure}
      \centering
      \textbf{b}
    \end{subfigure}
    \hspace{0.32\linewidth}
    \begin{subfigure}
      \centering
      \textbf{c}
    \end{subfigure}
    \newline
    \hfill
    \begin{subfigure}
      \centering
      \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Figure2a_PGcomcol4a_Bimp.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}
      \centering
      \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Figure2c_PGcomcol4a_Bimp.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}
      \centering
      \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Figure2e_PGcomcol4a_Bimp.png}
    \end{subfigure}
    \newline
    \begin{subfigure}
      \centering
      \textbf{d}
    \end{subfigure}
    \hspace{0.32\linewidth}
    \begin{subfigure}
      \centering
      \textbf{e}
    \end{subfigure}
    \hspace{0.32\linewidth}
    \begin{subfigure}
      \centering
      \textbf{f}
    \end{subfigure}
    \newline
    \hfill
    \begin{subfigure}
      \centering
      \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Figure2b_PGcomcol4a_Bimp.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}
      \centering
      \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Figure2d_PGcomcol4a_Bimp.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}
      \centering
      \includegraphics[width=0.3\linewidth,height=0.3\linewidth]{../Figures/Figure2f_PGcomcol4a_Bimp.png}
    \end{subfigure}
    \caption{Data quality report of forward reads from bumble bee sample \textit{Bee\_Bimpatiens\_hv4\_1}. 
    a, b and c shows low base quality in 3'-end, uneven base content in 5'-end and present of Nextera adaptors in raw data, respectively. 
    d, e and f shows the same results from clean data.}
    \label{QC}
    \end{figure}

  \subsection{Numbers of raw, clean and non-host read pairs}
  \begin{table}[H]
    \centering
    \caption{Statistics of sequenced read pairs.}
    \label{data_statistics}
    \resizebox{500pt}{!}{
    \begin{tabular}{@{}ccccc@{}}
    \toprule
    Sample                           &Host                     &Raw read pair   &\makecell[c]{Clean read pair\\\left[percentage\ of\ raw\ read\ pair\right]}&\makecell[c]{Non-host read pair\\\left[percentage\ of\ clean\ read\ pair]} \\ \midrule
    \textit{Bee\_Amellifera\_hv15\_1}&\textit{Apis mellifera}  &63159968&45059211 \left[71.34\%\right]&38837759 \left[86.19\%\right] \\
    \textit{Bee\_Amellifera\_wild\_1}&\textit{Apis mellifera}  &58113227&44466776 \left[76.52\%\right]&17014144 \left[38.26\%\right] \\
    \textit{Bee\_Amellifera\_hv13\_2}&\textit{Apis mellifera}  &56836899&35282101 \left[62.08\%\right]&10413704 \left[29.52\%\right] \\
    \textit{Bee\_Amellifera\_hv13\_1}&\textit{Apis mellifera}  &1104861&842095 \left[76.22\%\right]&665507 \left[79.03\%\right] \\
    \textit{Bee\_Bimpatiens\_hv3\_1} &\textit{Bombus impatiens}&63973750&53612702 \left[83.80\%\right]&5300592 \left[9.89\%\right] \\
    \textit{Bee\_Bimpatiens\_hv4\_1} &\textit{Bombus impatiens}&58988182&48426748 \left[82.10\%\right]&3557052 \left[7.35\%\right] \\
    \textit{Bee\_Bimpatiens\_hv4\_2}&\textit{Bombus impatiens} &54955553&45805759 \left[83.35\%\right]&4618023 \left[10.08\%\right] \\
    \textit{Flower\_eDNA}           &None                      &1443107&882436 \left[61.15\%\right]&882436 \left[100\%\right] \\
    \bottomrule
    \end{tabular}}
    \end{table}

  \subsection{Species composition of arthropods, plants, fungi and viruses}
  \begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{../Figures/RelativeAbundance_0_01_species_Arthropods.pdf}
    \caption{Heatmaps for arthropod species abundance distribution in all samples. 
    The relative abundance takes reads assigned to arthropod species as background. 
    Species with relative abundance smaller than 1\% in all samples are collapsed as "others". 
    It should be noted that for bee samples, host contamination was removed before taxon profiling. 
    As a result, the relative abundances of honey bees are extremely low in three honey bee samples, and the same for bumble bees in three bumble bee samples.}
    \label{ArthropodHeatmap}
    \end{figure}

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{../Figures/RelativeAbundance_0_01_species_Plants.pdf}
    \caption{Heatmaps for plant species abundance distribution in all samples. 
    The relative abundance takes reads assigned to plant species as background. 
    Species with relative abundance smaller than 1\% in all samples are collapsed as "others".}
    \label{PlantHeatmap}
    \end{figure}

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{../Figures/RelativeAbundance_0_01_species_Fungi.pdf}
    \caption{Heatmaps for fungal species abundance distribution in all samples. 
    The relative abundance takes reads assigned to fungal species as background. 
    Species with relative abundance smaller than 1\% in all samples are collapsed as "others".}
    \label{FungusHeatmap}
    \end{figure}

  \begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{../Figures/RelativeAbundance_0_01_species_Viruses.pdf}
    \caption{Heatmaps for virus species abundance distribution in all samples. 
    The relative abundance takes reads assigned to virus species as background. 
    Species with relative abundance smaller than 1\% in all samples are collapsed as "others".}
    \label{ViruseHeatmap}
    \end{figure}

  \subsection{Inferred pathways of carbohydrate and amino acid metabolism}
  \begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{../Figures/PathwayCov_Carbonhydrate.pdf}
    \caption{Heatmaps for pathways of carbohydrate metabolism.}
    \label{Carbohydrate}
    \end{figure}

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.8\textwidth]{../Figures/PathwayCov_AminoAcid.pdf}
      \caption{Heatmaps for pathways of amino acid metabolism.}
      \label{AminoAcid}
      \end{figure}

\end{document}